---
title: "Build test/train for round 2 model building"
author: "Eva Kleingeld"
date: "October 5, 2016"
output: html_document
---

First we read in a large dataset containing six days worth of GMS data. 
On four of these days temperatures were below freezing for (parts of) the day. 
On two days temperatures were always above freezing. 
The selected days each experienced different meteorological conditions.
For more details on the weather during the selected days we refer to the scripts:
- Met_conditions.Rmd
- Analyze_Days.R

```{r}
# Empty environment
rm(list=ls())

# Install packages if needed

# Read in the 6 Days data set
load("F:/KNMI/Six_Days.Rda")

# Rename dataframe
Six_Days <- Six_Days_3
rm(Six_Days_3)

# Examine data frame
str(Six_Days)
```

Next, we need to select a six hour *train* set and a 1 hour and 30 minute *test* set. 
The trains set will be used to build the model, the test set will be used to test it. 

We use the createDataPartition function from the caret package to make these sets. 

6 days times 24 hours is 144 hours. A subset of six hours from a total of 144 hours is 4.2% of all data. A subset of 1.5 hours is 1.05% of all data. 


```{r}
library(caret)



```







