---
title: "Lin_Reg"
author: "Eva Kleingeld"
date: "December 15, 2016"
output: html_document
---

First clear your working environment and install all necessary packages

```{r}
rm(list=ls())

# Install packages
#install.packages("caret")

```

Now load in the test and train sets and split them into target and predictor sets

```{r}
# Load train
load("/usr/people/kleingel/Projects/MLProject/Train_BIG_noCat.Rda")

# Load test
load("/usr/people/kleingel/Projects/MLProject/Test_BIG_noCat.Rda")

# Split train set into target and predictors 
Target_Train <- Train_set$TRoad
Train_set <- subset(Train_set, select=-c(TRoad))

# Split test set into target and predictors
Target_Test <- Test_set$TRoad
Test_set <- subset(Test_set, select = -c(TRoad))

```

# Build a linear model

To test:
Can you supply mutiple metrics? metric = c("RMSE", "Rsquared")

Possible cross-validation options:

cv = K-fold cross validation
repeatedcv = K-fold cross validation, repeated a certain nr. of times (why do this?)
LGOCV = leave group out cross validation 
boot = bootstrap
oob = out of bag (for Ranfom Forest)


```{r}
library(caret)
library(doParallel)
library(parallel)


#cluster_1<-makeCluster(7)
registerDoParallel(3)
#getDoParWorkers()


control<-trainControl(method="repeatedcv",repeats=5,number=10, allowParallel = TRUE) 
length <- 10

gbmGridReduced<-expand.grid(.shrinkage=c(0.1),.n.trees=3:5,.interaction.depth=4:6,.n.minobsinnode=3:5)

system.time(LinearModel <- train(x = Train_set, 
                                 y = Target_Train, 
                                 method = "lm", 
                                 metric = "RMSE",
                                 trControl = control,
                                 tuneLength = length))

summary(LinearModel)

LinearModel$results

# Predict
LM_Predict <- extractPrediction(models = list(LinearModel), testX = Test_set, testY = Target_Test)

# Plot observed versus predicted
plotObsVsPred(LM_Predict)


stopCluster(cluster)
registerDoSEQ()

#Save the model
save(LinearModel, file =  "/usr/people/kleingel/Projects/MLProject/LinearModel.Rda")

```

```{r}
library(caret)
library(doParallel)
library(parallel)
#library(foreach)
library(kernlab)

#cl<-makeCluster(7)
registerDoParallel(3)
#getDoParWorkers()

sigmaRangeReduced<-sigest(as.matrix(Target_Train))[1]
svmRadialRGridReduced<-expand.grid(C=2^(seq(-4,4)))

control_2<-trainControl(method="cv", number=10, allowParallel = TRUE) 

SVMradModel <- train(x  = Train_set, 
                     y = Target_Train, 
                     method = "svmRadialCost",
                     trControl = control_2)

                    # tuneGrid = svmRadialRGridReduced)
                    # ,
                    # trControl = control)


stopCluster(cluster)
registerDoSEQ()

# control<-trainControl(method="repeatedcv",repeats=5,number=10) 
# rfgridreduced <-expand.grid(mtry=2:3)
# length <- 10
# 
# RFModel <- train(x = Train_set, 
#                  y = Target_Train,
#                  method = "parRF",
#                  tuneGrid = rfgridreduced,
#                  importance = FALSE)
# 
# 
# 





```






