---
title: "Data Preprocessing including dummy variables"
author: "Eva Kleingeld"
date: "October 2, 2016"
output: pdf_document
---
 
In this script the test and train set are:
* subsetted so that they only include valid data
* split into test predictors/target variables and train predictors/target variables
* Altitude is heightened by 10m
* Temperatures are put in Kelvin instead of Celsius
* Data is preprocessed using the preProcess function, the transformations that are applied are:
+ Zero variance removal
+ BoxCox transform
+ centering
+ scaling
+ principal component analysis
* Next, some optional code is included for analyzing the remaining PC's 
* Finally, the resulting predictor sets and target variables are saved to .Rdata files

```{r}
# Empty environment
rm(list=ls())

# Install packages
# install.packages("corrplot")
# install.packages("GGally")
# install.packages("lubridate")

```

Read in the environmental dataset, train set and test set.
All sets are given a different name here and the old data frames are removed.
This way you can easily apply the script to different test/train sets.

The load_obj function was taken from this stackoverflow post:
[link](http://stackoverflow.com/questions/5577221/how-can-i-load-an-object-into-a-variable-name-that-i-specify-from-an-r-data-file)

```{r}
# Function to read in, rename an .Rdata file and remove the old .Rdata dataframe in one step
load_obj <- function(f)
{
    env <- new.env()
    nm <- load(f, env)[1]
    env[[nm]]
}

# Read in the environmental dataset
Env_Data <- load_obj("F:/KNMI/MLProject/Env_Data.Rda")
#Env_Data <- load_obj("/run/media/kleingel/Lexar/KNMI/Env_Data.Rda")
#Env_Data <- load_obj("/usr/people/kleingel/Projects/MLProject/Env_Data.Rda")

# Read in the train set
Train_set <- load_obj("F:/KNMI/Train_data_R2.Rda")
#Train_set <- load_obj("/run/media/kleingel/Lexar/KNMI/Train_data_R2.Rda")
#Train_set <- load_obj("/usr/people/kleingel/Projects/MLProject/Train_data_R2.Rda")


# Read in the test set
Test_set <- load_obj("F:/KNMI/Test_data_R2.Rda")
#Test_set <- load_obj("/run/media/kleingel/Lexar/KNMI/Test_data_R2.Rda")
#Test_set <- load_obj("/usr/people/kleingel/Projects/MLProject/Test_data_R2.Rda")

```

### Test the distribution of station numbers in the train and test sets

The histograms below show that the distribution of LOCATION (: stations) in the train and test sets is very similar. 
What is remarkable is that the distributions themselves are quite strange. You would expect all stations to occur equally often in the datasets. However, there seem to be a few 'peaks' in station numbers.
Is this because the data itself (:the original Six Days data) has a strange distribution of station numbers? Or is there a preferential selection of station numbers due to the stratified sampling?

We import the Six_Days data set below to test this. 

```{r}
# Test the distribution of station numbers in test and train

par(mfrow= c(1,2))

hist(Train_set$LOCATION)
hist(Test_set$LOCATION)

par(mfrow= c(1,1))

# Load the Six_Days dataset and plot LOCATION distribution next to train/test
# NOTE: This WILL make your computer run a lot more slowly
Six_Days <- load_obj("F:/KNMI/Six_Days.Rda")

par(mfrow= c(1,3))

hist(Six_Days$LOCATION, breaks = seq(1:1505),main = "Six_Days", freq = FALSE)
hist(Train_set$LOCATION, breaks = seq(1:1505), main = "Train",  freq = FALSE)
hist(Test_set$LOCATION, breaks = seq(1:1505), main = "Test",  freq = FALSE)

par(mfrow= c(1,1))

``` 


### Build the training set
```{r}
# Remove al data which is not labeled as 'valid' 
Train_set <- Train_set[Train_set$QUALITY == "valid", ]

# Drop the QUALITY column by name 
To_drop <- c("QUALITY")
Train_set <- Train_set[ ,!(names(Train_set) %in% To_drop)]

# Merge the train set and the environmental dataset
Train_set <-merge(Train_set,Env_Data,by.x=c("LOCATION","SENSOR"),by.y=c("MISD","SENSOR"))

# Split into input (predictors) and output (target variable)
# Target variable
Target_Train <- Train_set$TEMP

# Drop LOCATION/SENSOR/TIMESTAMP and TEMP
To_drop <- c("LOCATION", "SENSOR", "TIMESTAMP", "TEMP")
Train_set <- Train_set[ ,!(names(Train_set) %in% To_drop)]

# Place all predictors in one data frame
Predictors_Train <- Train_set

```

### Build the test set

```{r}
# Build the test set  -----------------------------------------------------

# Read in the test dataset and remove all data which is not labeled as 'valid' 
Test_set <- Test_set[Test_set$QUALITY == "valid", ]

# Drop the QUALITY column by name 
To_drop <- c("QUALITY")
Test_set <- Test_set[ ,!(names(Test_set) %in% To_drop)]

# Merge the test set and the environmental dataset
Test_set <-merge(Test_set,Env_Data,by.x=c("LOCATION","SENSOR"),by.y=c("MISD","SENSOR"))

# Split into input (predictors) and output (target variable)
# Output
Target_Test <- Test_set$TEMP

# Drop LOCATION/SENSOR/TIMESTAMP and TEMP
To_drop <- c("LOCATION", "SENSOR", "TIMESTAMP", "TEMP")
Test_set <- Test_set[ ,!(names(Test_set) %in% To_drop)]

# Place all predictors in one data frame
Predictors_Test <- Test_set

```

### Change data format 

Add 10m to all ALT values to ensure there are no negative heights. 
Change all temperatures from celsius to kelvin.

**Note: here we also change the target variables to kelvin, even though this is probably not necessary**   

```{r}
# Add 10m to ALT for train and test data
Train_set$ALT <- Train_set$ALT + 10
Test_set$ALT <- Test_set$ALT + 10

# Convert TD, TL in Predictors to Kelvin

# For train
Train_set$TL <- Train_set$TL + 273.15
Train_set$TD <- Train_set$TL + 273.15

# For test
Test_set$TL <- Test_set$TL + 273.15
Test_set$TD <- Test_set$TD + 273.15

# Convert target variables to Kelvin
Target_Train <- Target_Train + 273.15
Target_Test <- Target_Test + 273.15

# Save these sets of predictors 
# save(Train_set, file = "Train_set_unProcessed.Rda")
# save(Test_set, file = "Test_set_unProcessed.Rda")
```

# Test the distribution of predictors in test/train
There seems to be a slight difference in the distribution of stations in the train/test sets. See the difference in summary statistics calculated below. However, even though this difference is very minor, the PCA already goes wrong! 

```{r}

DiffSum <- function(Predictor){
  Summary1 <- summary(Predictors_Train[, Predictor])
  Summary2 <- summary(Predictors_Test[, Predictor])
  
  The_Diff <- Summary1 - Summary2
  
  return(The_Diff)
}

for (i in colnames(Predictors_Train)){
  print(i)
  print(DiffSum(i))
}



```

# PreProcessing

Here we remove variables with zero variance ("zv"), perform a BoxCox transform to resolve skewness ("BoxCox"), center and scale the data ("center", "scale") and perform a principal component analysis ("pca"). 

We remove zero variance predictors because these predictors cannot help the ML algorithms predict the target variable. Centering and scaling the data will speed up the model building process. A pca will reduce the number of predictors and will resolve multicollinearity, thereby speeding up the model building process. In order to perform a pca data needs to be centered and scaled (which we have already done) and the data much not be skewed. With a BoxCox transform we resolve skewness. 

Dummy variables are included in the pca.

```{r}
library(caret)

# Build the object containing the transform information
xTrans <- preProcess(Train_set, method = c("zv", "BoxCox",  "center", "scale", "pca"),
                     na.remove = TRUE)

# Summary of transform information
print(xTrans)

# Number of PC's 
xTrans$numComp

# BoxCox transformed predictors
xTrans$method$BoxCox


# Transform train set
Train_set <- predict(xTrans, Train_set)

# Transform test set
Test_set <- predict(xTrans, Test_set)

```

The total number of principal components is 'r xTrans$numComp'. As the summary of xTrans shows, all predictors were centered, scaled and pca transformed. Only 'r length(xTrans$method$BoxCox)' predictors were BoxCox transformed, namely: 'r xTrans$method$BoxCox'. 


# Analysis of PC's

This section of the code contains an analysis of the PC's in the test and train sets. 
If you wish to cut down the run time of the script this section of the code can be commented/not run without changing the test/train sets. 

First, we look at the variable loadings. 
The rotation column of the xTrans object stores the variable loadings. 
Each principal component after the PCA is a linear combination of the original predictors.The coefficient for each predictor is called loading. A variable loading close to 0 indicates that a predictor did not contribute much to the principal component

```{r}
xTrans$rotation

```

Next, a scree plot is produced which shows how much of the total variance is explained by the principal components. To build the scree plot I adapted a method from: [link](https://www.analyticsvidhya.com/blog/2016/03/practical-guide-principal-component-analysis-python/) 
Important to note!!! By applying the method to the PC's that you get after transformation you can only see the percentage variance explained for the PC's that came out of the preProcessing prediction. In other words: Only the PC's that together explain 95% of the variance are shown. Unfortunately, you cannot extract the rest of the PC's when you use caret to perform a PCA. 

```{r}
# Scree plot of the explained variance

# First, we build a function that makes scree plots of the PC's
Plot_Scree <-function(The_set){
  
  # Get the standard deviation per PC 
  PC_stdev <- apply(The_set, 2, sd, na.rm = TRUE)
  
  # Get the variance per PC by taking the square of the standard deviation per PC
  PC_var <- PC_stdev^2
  
  # Calculate proportion of total variance explained per PC by dividing PC_var by the total variance explained
  PC_prop_var <- PC_var/(sum(PC_var))
  
  # Make the scree plot
  plot(y = PC_prop_var, x = 1:length(The_set), type = "b", xlab = "Principal Components", ylab = "Proportion of variance explained")
}

# Make a scree plot of the train set PC's 
Plot_Scree(Train_set)

# Make a scree plot of the test set PC's 
Plot_Scree(Test_set)


```

To get an idea of what these PC's look like you can plot them in a pairwise plot (Also called a scatterplot matrix).

The PC's show that outliers may throw off your pca. 
For more information on this topic, see: 
[link](http://www.math.umn.edu/~lerman/Meetings/SIAM2012_Sujay.pdf) **Note: Do we need to solve this? (Probably yes, but...)** 

Here we make scatterplot matrices for the train set, but the same can be done for the test set.

**Note: Code takes a while to run, so is left commented. Uncomment to plot.**

```{r}
# Scatterplot matrices for the train set, 5x5 plots
#pairs(Train_set[, 1:5])
#pairs(Train_set[ , 6:10]) 
#pairs(Train_set[ , 11:15])
#pairs(Train_set[ , 16:20])
#pairs(Train_set[ , 20:25])

# The large plots stored in the plotting device can slow down your PC.
# In order to empty the plotting device type: dev.off()
```

Next, to check if the pca indeed removed multicollinearity a correlation plot is made. 
If the pca went well the PC's should not be correlated. 

```{r}
library(corrplot)

# Correlation plot for the train set
corTrain_1 <- cor(Train_set, use = "complete.obs")
corrplot(corTrain_1, method = "circle")

# Correlation plot for the test set
corTest_1 <- cor(Test_set, use = "complete.obs")
corrplot(corTest_1, method = "circle")

``` 

Next, we test for (near) zero variance in the test/train sets
```{r}
# Test for (near) zero variance in the train set
NearZero_Train <- nearZeroVar(Train_set, names = TRUE, saveMetrics = TRUE)
head(NearZero_Train)

# As expected, the are no zero variance variables, because those are removed during preProcessing
any(NearZero_Train$zeroVar)
# After pca there are also near zero variables
any(NearZero_Train$nzv)


# Test for (near) zero variance in the test set
NearZero_Test <- nearZeroVar(Test_set, names = TRUE, saveMetrics = TRUE)
head(NearZero_Test)

# As expected, the are no zero variance variables, because those are removed during preProcessing
any(NearZero_Test$zeroVar)
# After pca there are also near zero variables
any(NearZero_Test$nzv)

```

# Final step: Save sets

To use these sets in the scripts which build TW models we need to save the sets. 

```{r}
# The sets are saved to your working directory as Rdata files
# Your current working directory is:
getwd()
#setwd("C:/Users/Eva/Documents/Stage/ML_project")
setwd("/usr/people/kleingel/Projects/MLProject")

# Save the test/train predictor sets 
save(Train_set, file = "Train_set_R2.Rda")
save(Test_set, file = "Test_set_R2.Rda")

# Save the corresponding target variables
save(Target_Train, file = "Target_train_R2.Rda")
save(Target_Test, file = "Target_test_R2.Rda")

```
