---
title: "Testing the preProcess function"
author: "E Kleingeld"
output: html_document
---


This script is used to test how the preProcess function from the caret package transforms a set of predictors when you apply several different methods

### Read in the predictors

First we read in a set of predictors 
```{r}
# Empty environment
rm(list=ls())

# Read in the predictors
load("C:/Users/Eva/Documents/Stage/ML_project/Predictors_Test_6h.Rda")


```
This is what the data looks like: 

```{r echo=FALSE}
head(Predictors_Test)
``` 


### Analyse data before applying the preProcess function 

First, plot the correlation between all of the continuous valriables:
```{r}
library(corrplot)

corTest <- cor(Predictors_Test[, 1:6], use = "complete.obs")
corrplot(corTest, method = "number")

```

Next, plot the raw data

```{r} 
Cont_vars <- colnames(Predictors_Test)[1:6]

# Get a summary for each variable in the train data
for (i in seq_along(Cont_vars)){
    plot(Predictors_Test[ , i], ylab = Cont_vars[i])
}

````

### Test removal of zero variance and BoxCox

Below we apply the preProcess function so that all zero variance predictors are removed and so that a BoxCox transform is applied. The BoxCox transform should unskew the data. This is necessary because a PCA fails when you apply it on skewed data. The most skewed variable is altitude. 

After transforming with BoxCox, altitude is no longer skewed. What is surprising is that TL and TD are not transformed in the same way: TL becomes very small whilst TD becomes very large. ** why does this happen? ** 

After the BoxCox transform the correlations between the predictors is identical to that before the transform. 

```{r}
library(caret)
xTrans <- preProcess(Predictors_Test[, 1:6], method = c("zv", "BoxCox"),
                     na.remove = TRUE)

Predictors_BoxCox <- predict(xTrans, Predictors_Test[ ,1:6])

# Plot each of the BoxCox transformed predictors
Cont_vars <- colnames(Predictors_BoxCox)

for (i in seq_along(Cont_vars)){
    plot(Predictors_BoxCox[ , i], ylab = Cont_vars[i])
}

# Test the correlations between the BoxCox transformed predictors
corBoxCox <- cor(Predictors_BoxCox, use = "complete.obs")
corrplot(corBoxCox, method = "number")

```

### Test removal of zero variance and BoxCox + centering and scaling
When we center the data the mean of the predictor's data is subtracted from the predictor values. 
When we scale the data the predictor values are divided by the standard deviation.

When you center and scale predictors the predictor values vary over a smaller range. For some ML algorithms, such as neural networks or support vector machines, this is especially useful because they will run faster on scaled/centered data. 


After the BoxCox transform and centering/scaling the correlations between the predictors is identical to that before the transform. 


```{r}
xTrans <- preProcess(Predictors_Test[, 1:6], method = c("zv", "BoxCox",  "center", "scale"),
                     na.remove = TRUE)

Predictors_centerScale <- predict(xTrans, Predictors_Test[ ,1:6])

# Plot each of the transformed variables
Cont_vars <- colnames(Predictors_centerScale)

for (i in seq_along(Cont_vars)){
    plot(Predictors_centerScale[ , i], ylab = Cont_vars[i])
}
# Test the correlations between the BoxCox + centered/scaled transformed predictors
corBoxCox_CS <- cor(Predictors_centerScale, use = "complete.obs")
corrplot(corBoxCox_CS, method = "number")
```

### Test removal of zero variance and BoxCox + centering and scaling + PCA
```{r}
xTrans <- preProcess(Predictors_Test[, 1:6], method = c("zv", "BoxCox",  "center", "scale", "pca"),
                     na.remove = TRUE)

Predictors_PCA <- predict(xTrans, Predictors_Test[ ,1:6])

# Plot each of the transformed variables
Cont_vars <- colnames(Predictors_PCA)

for (i in seq_along(Cont_vars)){
    plot(Predictors_PCA[ , i], ylab = Cont_vars[i])
}
# Test the correlations between the BoxCox + centered/scaled transformed predictors
corBoxCox_CS_PCA <- cor(Predictors_PCA, use = "complete.obs")
corrplot(corBoxCox_CS_PCA, method = "number")
```



